\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png}\\ \vspace{5mm} LFEUI - Logbook 2\vspace{-5mm}

### 2024-01-24 text/LogBooks/logbook2.tex
Wednesday, 24. January 2024 at 16:05:01

First of all, we assembled the source holder like it is described in the previous section. So that we can do a calibration from the number of counts per channel to energy, we put an additional americium-241 source on the holder.
Using a ruler, we measured the distance between the samples. We put the sample holder on a support and afterwards both of them inside a chamber. A vacuum machine was connected to the accelerator to reduce the pressure inside the accelerator.

\begin{table}[h!]
\centering
\begin{tabular}{|w{2cm}|w{4cm}|}
\hline
Samples & Distance [cm]  \\ \hline
Americium & 1.5  \\ \hline
LiF & 3.3 \\ \hline
LiAlO2 & 4.8  \\ \hline
Implanted Li-7 in Al & 6.0  \\ \hline
\end{tabular}
\caption{Distance between the samples in the sample holder}
\label{tab:samples}
\end{table}

Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the manuals provided. Firstly, we turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses. Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.

In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples. Then we turned on the camera next to the samples to observe if it was well adjusted. The beam wasn't rightly centred so we adjusted the angles and positions of the lenses until it was in the correct position. Finally, we used a collimator with a 2mm radius to collimate the beam.

We used a computer program to obtain the spectra of the different sources. To start, we obtained the spectrum of americium-241 in order to do a calibration posteriorly. Then we obtained the spectra of the different sources. The multichannel analyzer (MCA) utilized to acquire data was gradually set to 70 V and the angle between the top two detectors was 165º.

After obtaining the lithium spectra, we obtained the spectrum of borum as well. Borum can be detected in the middle of the sources since the encapsulation of the sources is made of boric acid.

eletromagnets adjustment: there are 3 magnets:
- low energy switch magnet current - 1.7 A;
- high energy switch 90º magnet current - 13.2 A;
- second high energy switch magnet current - 14.4 A;
1315 keV

proton source adjustment - 15 A

Temos um feiexe e conseguimos ver. ajustamos o feixe com os campos magenticos e com o steering e com o electrostatic steerer no fim. Depois colimamos o nosso feixe através de metodos oftalmologicos. (colimador de raio 2mm)

Eletronica de aquisicao de dados a 70V, evoluindo gradualmente.O angulo entre os detetores é 165º (valor q usamos no calculo da energia)
Offset porta alvos - 15mm
Escrever a distancia entre as amostras no alvo

Virar po porta alvos de costas para ter o americio241 de frente de modo a calibrar

Desviar a amostra 18 mm de modo a ter a 1 amostra a contar de cima. Mudar o modo de aquisição de tempo para carga (tempo para a fonte radioativa para calibrar e carga para as fontes com litio para o detetar)
O encapsulamento das amostras é feito com acido borico pelo que a detecao de boro é no meio das amostras (pastilhas).

We left CTN at around 16:45 PM.

>>>>

First of all, we assembled the source holder like it is described in the previous section. So that we can do a calibration from the number of counts per channel to energy, we put an additional americium-241 source on the holder.
Using a pachymeter, we measured the distance between the samples. We put the sample holder on a support and afterwards both of them inside the chamber. The vacuum machine was connected to the accelerator to reduce the pressure inside the accelerator (measured from the top of the container).

\begin{table}[h!]
\centering
\begin{tabular}{|w{4cm}|w{3cm}|}
\hline
Samples & Distance [cm]  \\ \hline
Americium & 1.5  \\ \hline
LiF & 3.3 \\ \hline
LiAlO2 & 4.8  \\ \hline
Implanted Li-7 in Al & 6.0  \\ \hline
\end{tabular}
\caption{Distance between the samples in the sample holder}
\label{tab:samples}
\end{table}

Afterwards, we started the procedure to turn on the TANDEM accelerator, according to the manuals provided. This was a very time consuming step that took all morning. We had to start the high voltage supply, start the vacuum in the first chamber of the accelerator and then extend it to the accelerator as a whole and gradually increase all the accelerator parameters to the standards used in previous logbooks.

In the afternoon, we fisrtly turned on the $\ce{H^+}$ duoplasmatron source of the accelerator as well as the magnets and lenses. Then, we opened the valves to propagate the vacuum throughout the entirety of the tubes.

In the control panel, we adjusted the angles of the lenses and the magnets current according to the manual, so that the beam would reach the samples.

The electromagnets were set with the following currents:
- low energy switch magnet current - 1.5 A;
- high energy switch 90º magnet current - 13.1 A;
- second high energy switch magnet current - 13.9 A;

Then we turned on the camera next to the samples to observe if it was well adjusted. The beam wasn't rightly centred so we adjusted the angles and positions of the lenses until it was in the correct position using electrostatic steering. Finally, we used a collimator with a 2mm radius to collimate the beam.

The angle between the top two detectors was 165º, a value used in the proposal to get an estimate for the energy used for the beam. The final energy reached by the tandem accelerator was 650 keV, or 1300 keV total (as the tandem accelerator has 2 tubes).

We used a computer program to obtain the spectra of the different sources. To start, we obtained the spectrum of americium-241 in order to do a calibration subsequently. Then we changed the acquisition mode from time to charge to obtain the spectra of the different sources, as the time mode is only used for calibrations. The multichannel analyser (MCA) utilised to acquire data was gradually set to 70 V.

After obtaining the lithium spectra, we obtained a spectrum with borum by aiming the detector at the middle of the encapsulated sources (approximately 5.4 cm from the top), since their encapsulation is made of boric acid.

We obtained 8 files with our spectra, all with 1024 channels: from TT\_20 to TT\_27. In those files we were only interested in the unfiltered spectra of channels 0, 1 and 2 - channel 2 corresponded to the detector under the sample holder.

TT\_20 was a very short calibration that shall not be used.

TT\_21 was the used calibration and it took 5 minutes to acquire.

TT\_22 was our first acquisition of the top sample (LiF). Max charge was set to 2 micro Coulombs.

TT\_23 was the acquisition of the same sample but with with a max charge of 2 micro Coulombs.

TT\_24 was the acquisition of LiAlO2 with a max charge of 2 micro Coulombs.

TT\_25 had the spectrum of the implanted sample with a 2 micro Coulombs max charge.

TT\_26 was the same sample but with a max charge of 20 micro Coulombs as the implanted lithium was had very small peaks.

TT\_27 we were aiming at the middle of the sources to obtain a spectrum with borum. Max charge of 20 micro Coulombs.

At the end of the acquisition we had to gradually reduce the current and voltage of the accelerator in order to turn it off. Basically doing the inverse procedure we did in the morning.

We left CTN at around 4:45 PM.

\vspace{3cm}
